\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2019-08-24 icse20/intro.tex
temporal update

\section{Introduction}
JavaScript is one of the widely used programming languages not only for client-side
but also for server-side programming~\cite{nodejs, meanjs}
and even for small embedded systems~\cite{espruino, tessel2}.
It is the top-ranked language used in active GitHub
repositories\footnote{https://githut.info/}, and \#7 in the TIOBE
Programming Community index\footnote{https://www.tiobe.com/tiobe-index/}.
According to W3Techs\footnote{https://w3techs.com/technologies/details/cp-javascript/all/all},
95.2\% of websites use JavaScript as their client-side programming language.

Despite its popularity, JavaScript developers often suffer from its intricate semantics,
which may cause security vulnerabilities.  For example, calling the following JavaScript
function seems to always return \( \code{false} \):
\begin{lstlisting}[style=myJSstyle]
    function f(x) { return x == !x; }
\end{lstlisting}
Unfortunately, it returns \( \code{true} \) when its argument is an empty array
\( \code{[]} \).  To correctly understand and reason about such complex
behaviors, the formal semantics of JavaScript is necessary.

@@

Moreover, manual extraction does not scale for frequent updates of the
specification.  While most existing formal semantics and analyzers of
JavaScript support only ECMAScript 5.1~\cite{ecma5} released in December 2009,
from ECMAScript 6, officially dubbed ECMAScript 2015, the 
ECMAScript specification is updated every year.  Focusing on
ECMAScript 5.1 was reasonable until 2014, because the
specification was rarely updated.  However, in late 2014, The Ecma
Technical Committee 39 (TC39)~\cite{tc39} decided to release the
specification annually.  Currently, no formal semantics
nor analyzers for JavaScript correctly deal with recent ECMAScript
features such as lexical binding via \( \code{let} \), the spread
\( \code{...} \) operator, classes, the \( \code{for-of} \) operator,
the \( \code{async} \) functions, and generators.

>>>>

\section{Introduction}
JavaScript is one of the most widely used programming languages not only for client-side
but also for server-side programming~\cite{nodejs, meanjs}
and even for small embedded systems~\cite{espruino, tessel2}.
It is the top-ranked language used in active GitHub
repositories\footnote{https://githut.info/}, and \#7 in the TIOBE
Programming Community index\footnote{https://www.tiobe.com/tiobe-index/}.
According to W3Techs\footnote{https://w3techs.com/technologies/details/cp-javascript/all/all},
95.2\% of websites use JavaScript as their client-side programming language.

Despite its popularity, JavaScript developers often suffer from its intricate semantics,
which may cause unexpected behavior.  For example, calling the following JavaScript
function seems to always return \( \code{false} \):
\begin{lstlisting}[style=myJSstyle]
    function f(x) { return x == !x; }
\end{lstlisting}
Unfortunately, it returns \( \code{true} \) when its argument is an empty array
\( \code{[]} \).  To correctly understand and reason about such complex
behavior, the formal semantics of JavaScript is necessary.

@@

Moreover, manual extraction does not scale for frequent updates of the
specification.  Most existing formal semantics and analyzers of
JavaScript support only ECMAScript 5.1~\cite{ecma5} released in June 2011.
Focusing on ECMAScript 5.1 was reasonable until 2014, because the
specification was rarely updated.  However, in late 2014, The Ecma
Technical Committee 39 (TC39)~\cite{tc39} decided to release the
specification annually.  Currently, no formal semantics
nor analyzers for JavaScript correctly deal with recent ECMAScript
features such as lexical binding via \( \code{let} \), the spread
\( \code{...} \) operator, classes, the \( \code{for-of} \) operator,
the \( \code{async} \) functions, and generators.

### 2019-08-24 icse20/main.tex
temporal update

To alleviate the problem, we propose \( \tool \), a JavaScript IR-based Semantics
Extraction Toolchain.  It \textit{automatically} extracts the JavaScript syntax and
semantics from ECMAScript specifications.  \( \tool \) generates a JavaScript
parser for a given grammar written in \( \bnfes \), an extended BNF
used in ECMAScript specifications.  It also converts each
abstract algorithm written in a structured natural language into
functions written in \( \ires \), an Intermediate Representation
for ECMAScript specifications.  The compilation from algorithms to
\( \ires \) functions is defined with \textit{compile rules}.
We applied \( \tool \) to ECMAScript 2020,
and the automatically extracted semantics passed 19,220 of 19,282
applicable tests in Test262, the official ECMAScript conformance test
suite.  The tool also extracted the semantics of an incomplete proposal,
and it detected three errors from the specification and they were all
confirmed.
\end{abstract}

\keywords{
  JavaScript,
  mechanized formal semantics,
  recursive descent parsing
}

>>>>

To alleviate the problem, we propose \( \tool \), a JavaScript IR-based Semantics
Extraction Toolchain. It \textit{synthesizes} a JavaScript
parser for a given grammar written in \( \bnfes \), an extended BNF
used in ECMAScript. It also \textit{automatically} extracts
semantics based on rule-based compilations. Each \textit{compile rule}
describes how to convert each abstract algorithm written in a structured natural language
into functions written in \( \ires \), an Intermediate Representation
for ECMAScript. \( \tool \) successfully synthesizes
parsers and compiles 91.60\% of algorithm steps for five recent
ECMAScript versions. We tested the extracted semantics from the next version
ECMAScript 2020 using Test262, the official ECMAScript conformance test
suite. It passed 19,220 of 19,282 tests and we found three specification errors
using our tool. Moreover, we check the adaptability of \( \tool \)
via a case study for an incomplete proposal.
\end{abstract}

\keywords{
  JavaScript,
  mechanized formal semantics,
  program synthesis
}

### 2019-08-24 icse20/overview.tex
temporal update

\textbf{Syntax.} The ECMAScript specification provides the lexical and
syntactic grammars in Appendix A using their own extended Backus-Naur
Form (BNF) for ECMAScript.  We dub it \( \bnfes \) and formally define
it in Section~\ref{sec:parser}.  Our {\sf Spec Extractor} reads the
grammars written in \( \bnfes \) and converts them into JSON files.
For example, Figure~\ref{fig:array-literal}(a) shows the
\textit{ArrayLiteral} production in the ECMAScript 2020
specification.  It takes two boolean parameters \textsf{Yield} and
\textsf{Await} and has three alternatives.  The first alternative
consists of three symbols: two terminal symbols \( \code{[} \) and
\( \code{]} \) and one non-terminal symbol \textit{Elision}$_{\mbox{\footnotesize opt}}$.
The {\small opt} subscript denotes that it is optional.  In the second
and third alternatives, \textit{ElementList}$_{\mbox{\footnotesize\sf [?Yield, ?Await]}}$
denotes a parametric non-terminal symbol \textit{ElementList}
with the parameters \textsf{Yield} and \textsf{Await} of
\textit{ArrayLiteral} as its two arguments. The prefix \textsf{\small ?}
of a symbol denotes that the symbol is passed as an argument.

@@

Finally, \( \tool \) constructs {\sf AST-\( \ires \) Translator} with
the given \( \ires \) functions and \textit{manually} specified
{\sf Global Setting}, which have minor but necessary information to
evaluate JavaScript programs described in the ECMAScript specification.
They consists of the structure of the standard built-in
objects and ECMAScript data types.

>>>>

\textbf{Syntax.} The ECMAScript specification provides the lexical and
syntactic grammars in Appendix A using their own extended Backus-Naur
Form (BNF) for ECMAScript.  We dub it \( \bnfes \) and formally define
it in Section~\ref{sec:parser}.  Our {\sf Spec Extractor} reads the
grammars written in \( \bnfes \) and converts them into JSON files.
For example, Figure~\ref{fig:array-literal}(a) shows the
\textit{ArrayLiteral} production in the ECMAScript 2020
specification.  It takes two boolean parameters \textsf{Yield} and
\textsf{Await} and has three alternatives.  The first alternative
consists of three symbols: two terminal symbols \( \code{[} \) and
\( \code{]} \), and one non-terminal symbol \textit{Elision}$_{\mbox{\footnotesize opt}}$.
The {\small opt} subscript denotes that it is optional.  In the second
and third alternatives, \textit{ElementList}$_{\mbox{\footnotesize\sf [?Yield, ?Await]}}$
denotes a parametric non-terminal symbol \textit{ElementList}
with the parameters \textsf{Yield} and \textsf{Await} of
\textit{ArrayLiteral} as its two arguments. The prefix \textsf{\small ?}
of a symbol denotes that the symbol is passed as an argument.

@@

Finally, \( \tool \) constructs {\sf AST-\( \ires \) Translator} with
the given \( \ires \) functions and \textit{manually} specified
{\sf Global Setting}, which has minor but necessary information to
evaluate JavaScript programs described in the ECMAScript specification.
They consists of the structure of the standard built-in
objects and ECMAScript data types.
